---
title: "HW2 Part C"
author: "Omair Shafi Ahmed"
date: "10/10/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Importing Data
```{r}
if(!require('ggplot2')) install.packages("ggplot2",repos = "http://cran.us.r-project.org")
if(!require('dplyr')) install.packages("dplyr",repos = "http://cran.us.r-project.org")
if(!require('RMySQL')) install.packages("RMySQL",repos = "http://cran.us.r-project.org")
if(!require('dbplyr')) install.packages("dbplyr",repos = "http://cran.us.r-project.org")
if(!require('rstudioapi')) install.packages("rstudioapi",repos = "http://cran.us.r-project.org")



library('ggplot2')
library('dplyr')
library ('RMySQL')
library('dbplyr')
library('rstudioapi')
```

Connecting to DB

```{r}

con <- DBI::dbConnect(RMySQL::MySQL(), 
  host = "localhost",
  socket = "/Applications/MAMP/tmp/mysql/mysql.sock",
  port= 8889,
  user = "root",
  #password = rstudioapi::askForPassword("Database password")
  password = "root",
  db = "DBLP"
)

authors <- tbl(con, "authors")
general <- tbl(con, "general")

```


## Problem 5

Filter the data to include only the authors for whom a gender was predicted with a probability of 0.95 or greater, and then create a bar plot showing the number of distinct male and female authors in the dataset.

```{r}



filtered_authors <- authors %>% filter(prob > 0.95, prob < 1) %>%
                    collect()

filtered_authors %>%
  group_by(gender) %>%
  count() %>%
  ggplot() +
    geom_col(mapping = aes(x= gender, y= n)) +
    xlab("Gender") +
    ylab("Papers published")

```
It appears as if males authors generally dominate in numbers of papers published.


##Problem 6
Again including only the authors for whom a gender was predicted with a probability of 0.95 or greater, create a stacked bar plot showing the number of distinct male and female authors published each year.


```{r}

filtered_authors %>%
  inner_join(general, key = 'k', copy=TRUE) %>%
  select(gender, year) %>%
  group_by(gender, year) %>%
  count() %>%
  ggplot() +
    geom_col(mapping = aes(x = year, y = n, fill = gender)) +
    xlab("Year") +
    ylab("Papers published")

```

The proportion of female scientists publishing papers have increased in recent years

## Problem 7

```{r}
filtered_authors %>%
  inner_join(general, key = 'k', copy = TRUE) %>%
  select(gender, year) %>%
  group_by(year) %>%
  summarise(`Male` = mean(gender == "M"),
            `Female` = mean(gender == "F")) %>%
  gather(-year, key = "Gender", value = "val") %>%
  ggplot() +
    geom_col(mapping = aes(x = year, y = val, fill = Gender)) +
    xlab("Year") +
    ylab("Gender Proportions")
```

Here, the increase in the proportion of female authors publishing papers is clearly visible.


